\thetitle

### 2020-11-22 sun_phases/spin_qudit_tomo.tex
minor changes to the main text

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}
\renewcommand{\d}{\text{d}}

@@

\newcommand{\ZZ}{\mathbb{Z}}

@@

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

@@

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

@@

  \S_V^2 \equiv \sum_k\f1{\p{M^V_k}^2} = \norm{M_V^{-1}}^2,

@@

is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, satisfying $M_V M_V^{-1}=1$.

@@

Note that the classical error scale $\S_V$ diverges if the measurement matrix $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V$ and finding estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$-sized blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular value decomposition takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing the classical error scale $\S_V$ takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{qudit_times.pdf}
  \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, or $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
    Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$ [\red{todo: update/correct after more simulations}].
    These results do not count fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of $V$ or $\rho$.
    The simulations used to benchmark runtime found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the constraints that all $\bk{\Pi_{\v v\mu}}_\rho\ge0$ and $\tr\p\rho=1$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2 < \f{\epsilon_V^2}{n},
  &&
  \epsilon_V^2 \equiv \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:bound}
\end{align}
where the {\it quantum error scale} $\epsilon_V$, so called because it is set by quantum shot noise, is defined in terms of (half of) the spectral range of $T_\ell$:
\begin{align}
  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}.
\end{align}
If $d$ is even or $\ell$ is odd, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
For comparison with previous error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V$ is the same as that of $\S_V$ (see Figure \ref{fig:times}).

Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact_error}, can be written in the form
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
  \label{eq:error}
\end{align}
where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated after constructing $\tilde\rho_V$ by computing $\E_V\p{\tilde\rho_V}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Revisiting the reconstruction error bound}
\label{sec:bound}

We now revisit the derivation of reconstruction error in Section \ref{sec:error} to make use of the bounds on variances $\bbk{\epsilon_{i\ell}^2}$.
To recap, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
\end{align}
which can be block diagonalized as
\begin{align}
  U^\dag M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  U \equiv \sum_{\ell,m} \oket{T_{\ell m}} \bra{\ell m},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v},
\end{align}
where $D^\ell_{mn}\p{\v v}\equiv\bk{\ell m|R\p{\v v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots )$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
These singular vectors and values define the orthonormal operators
\begin{align}
  Q^V_{\ell m} \equiv \sum_i q^V_{\ell mi} T_{i\ell},
  &&
  \v q^V_{\ell m} \equiv \f{\v x^V_{\ell m}}{M^V_{\ell m}},
\end{align}
where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$.
The state $\rho$ can be expanded in the basis of these operators as
\begin{align}
  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m},
\end{align}
and the estimates $\tilde T_{i\ell}$ of $\bk{T_{i\ell}}_\rho$ can be used to construct the following estimate $\tilde\rho_V$ of $\rho$:
\begin{align}
  \tilde\rho_V \equiv \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}} Q^V_{\ell m}
  \approx \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho} Q^V_{\ell m}
  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m}
  = \rho.
\end{align}
Recalling that $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$, we can use orthonormality of all $Q^V_{\ell m}$ to expand the mean squared distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2 \equiv \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
  < \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:bound_apndx}
\end{align}
where we used the fact that $\bbk{\epsilon_{i\ell}^2}\le\Gamma_\ell^2/n$, and
\begin{align}
  \sum_{m,i} \abs{q^V_{\ell mi}}^2
  = \sum_m \f1{\p{M^V_{\ell m}}^2}
  = \norm{M_{V\ell}^{-1}}
  = \S_{V\ell}^2.
\end{align}
The fact that $\bbk{\epsilon_{i\ell}^2}<1/2n$ also implies that
\begin{align}
  \E_V\p{\rho}^2 < \f1{2n} \sum_\ell \S_{V\ell}^2 = \f{\S_V^2}{2n}.
\end{align}
Note that the bound in Eq.~\eqref{eq:bound_apndx} is not tight, as the individual bounds on the variances $\bbk{\epsilon_{i\ell}^2}$ cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.

@@

Here we find exact expressions for reconstruction error, which can be used to estimate the error in a given reconstruction $\tilde\rho_V$ of an unknown state $\rho$ after performing tomography.
To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:bound} to write
\begin{align}
  \E_V\p{\rho}^2
  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell}\epsilon_{i'\ell}}
  = \f1n \sum_{\ell,m,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
  &&
  \abs{\tilde{\v q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2,
  \label{eq:error_SM_start}
\end{align}
where $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
We can therefore simplify
\begin{align}
  \abs{\tilde{\v q}_{\ell i}}^2
  = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
  = \bk{\v v_i| M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag | \v v_i}.
\end{align}
Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can also expand
\begin{align}
  \cov_\rho\p{T_{i\ell},T_{i\ell}}
  = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
  = \sum_{m,m'} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
\end{align}
which implies that
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_{i,\ell,m,m'}
  D^\ell_{m',0}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
  D^\ell_{m,0}\p{\v v_i}^*
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.
\end{align}
Altogether, this reconstruction error can be expressed more compactly by defining the {\it covariance matrix}
\begin{align}
  \C_\ell\sp{\rho} \equiv
  \sum_{m,m'} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}} \op{m}{m'},
\end{align}
and the {\it noise matrix}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, in terms of which
\begin{align}
  \E_V\p{\rho}^2 = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}},
  \label{eq:error_SM_mid}
\end{align}
where $\obk{X|Y}=\tr\p{X^\dag Y}$ is a trace inner product.

>>>>

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

@@

@@

@@

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering

@@

  \S_V^2 \equiv \sum_k \p{M^V_k}^{-2} = \norm{M_V^{-1}}^2,

@@

where $M_V^{-1}$ is the right inverse of $M_V$, satisfying $M_V M_V^{-1}=1$.

@@

Note that the classical error scale $\S_V$ diverges if the measurement matrix $M_V$ is singular, which indicates that measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V$ and estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$-sized blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular value decomposition takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing the classical error scale $\S_V$ takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{qudit_times.pdf}
  \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, or $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
    Each point is an average over $10^3$ calculations or 5 minutes of runtime, whichever comes first.
    These results do not count fixed runtimes to pre-compute quantities that can be recycled for every new choice of $V$ or $\rho$.
    Dashed lines show fits to a runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These simulations found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the constraints that $\tr\p\rho=1$ and $\bk{\Pi_{\v v\mu}}_\rho\ge0$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2 < \f{\epsilon_V^2}{n},
  &&
  \epsilon_V^2 \equiv \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:bound}
\end{align}
where the {\it quantum error scale} $\epsilon_V$, so called because it is set by quantum shot noise, is defined in terms of (half of) the spectral range of $T_\ell$:
\begin{align}
  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}.
\end{align}
If $d$ is even or $\ell$ is odd, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
For comparison with the ``classical'' error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V$ is the same as that of $\S_V$ (see Figure \ref{fig:times}).

Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$ by $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact_error}, can be written in the form
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
  \label{eq:error}
\end{align}
where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Revisiting the reconstruction error bound}
\label{sec:revisiting}

We now revisit the derivation of reconstruction error in Section \ref{sec:error} to make use of the bounds on variances $\bbk{\epsilon_{i\ell}^2}$.
To recap, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
\end{align}
which can be block diagonalized as
\begin{align}
  U^\dag M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  U \equiv \sum_{\ell,m} \oket{T_{\ell m}} \bra{\ell m},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v},
\end{align}
where $D^\ell_{mn}\p{\v v}\equiv\bk{\ell m|R\p{\v v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots )$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
These singular vectors and values define the orthonormal operators
\begin{align}
  Q^V_{\ell m} \equiv \sum_i q^V_{\ell mi} T_{i\ell},
  &&
  \v q^V_{\ell m} \equiv \f{\v x^V_{\ell m}}{M^V_{\ell m}},
\end{align}
where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$.
The state $\rho$ can be expanded in the basis of these operators as
\begin{align}
  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m},
\end{align}
and the estimates $\tilde T_{i\ell}$ of $\bk{T_{i\ell}}_\rho$ can be used to construct the following estimate $\tilde\rho_V$ of $\rho$:
\begin{align}
  \tilde\rho_V \equiv \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}} Q^V_{\ell m}
  \approx \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho} Q^V_{\ell m}
  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m}
  = \rho.
\end{align}
Recalling that $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$, we can use orthonormality of all $Q^V_{\ell m}$ to expand the mean squared distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2 \equiv \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
  < \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:bound_apndx}
\end{align}
where we used the fact that $\bbk{\epsilon_{i\ell}^2}\le\Gamma_\ell^2/n$, and
\begin{align}
  \sum_{m,i} \abs{q^V_{\ell mi}}^2
  = \sum_m \p{M^V_{\ell m}}^{-2}
  = \norm{M_{V\ell}^{-1}}
  = \S_{V\ell}^2.
\end{align}
The fact that $\bbk{\epsilon_{i\ell}^2}<1/2n$ also implies that
\begin{align}
  \E_V\p{\rho}^2 < \f1{2n} \sum_\ell \S_{V\ell}^2 = \f{\S_V^2}{2n}.
\end{align}
Note that the bound in Eq.~\eqref{eq:bound_apndx} is not tight, as the individual bounds on the variances $\bbk{\epsilon_{i\ell}^2}$ cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.

@@

Here we find exact expressions for reconstruction error, which can be used to estimate the error in a given reconstruction $\tilde\rho_V$ of an unknown state $\rho$ after performing tomography.
To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:revisiting} to write
\begin{align}
  \E_V\p{\rho}^2
  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell}\epsilon_{i'\ell}}
  = \f1n \sum_{\ell,m,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
  &&
  \abs{\tilde{\v q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2,
  \label{eq:error_SM_start}
\end{align}
where $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
We can therefore simplify
\begin{align}
  \abs{\tilde{\v q}_{\ell i}}^2
  = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
  = \bk{\v v_i| M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag | \v v_i}.
\end{align}
Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can also expand
\begin{align}
  \cov_\rho\p{T_{i\ell},T_{i\ell}}
  = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
  = \sum_{m,m'} D^\ell_m\p{\v v_i}^* D^\ell_{m'}\p{\v v_i}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
\end{align}
which implies that
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_{i,\ell,m,m'}
  D^\ell_{m'}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
  D^\ell_m\p{\v v_i}^*
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.
\end{align}
Altogether, this reconstruction error can be expressed more compactly by defining the {\it covariance matrix}
\begin{align}
  \C_\ell\sp{\rho} \equiv
  \sum_{m,m'} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}} \op{m}{m'},
\end{align}
and the {\it noise matrix}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, in terms of which
\begin{align}
  \E_V\p{\rho}^2 = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}},
  \label{eq:error_SM_mid}
\end{align}
where $\obk{X|Y}=\tr\p{X^\dag Y}$ is a trace inner product.
